---
title: "Regression Script to Test for Relationship with Physiology"
output: html_notebook
---

Set up and produce dataframes
```{r}
library(qiime2R)
library(tidyverse)
library(phyloseq)

# import rarefied feature table for climate chamber lizards only 
phy_clim<-qiime2R::qza_to_phyloseq("./Climate-chamber/core-metrics-results/rarefied_table.qza", "rooted-tree.qza","silva-taxonomy.qza", "physiologyafterexperiment.txt")

## convert to relative abundances
phy_sourcepop_norm<-transform_sample_counts(phy_clim, function(x) x/sum(x))

#agglomerate taxa at phylum -  on the count of those OTUs in each sample. 
glom<-tax_glom(phy_sourcepop_norm, taxrank = "Phylum")

#convert to dataframe
dataf<- psmelt(glom)

#convert phylum names to character vectors
dataf$Phylum<-as.character(dataf$Phylum)

#filter for only that phylum 
Firmicutes<-filter(dataf, Phylum == "Firmicutes")

Bacteroidetes<-filter(dataf, Phylum == "Bacteroidetes")

Proteobacteria<-filter(dataf, Phylum == "Proteobacteria")

Verrucomicrobia<-filter(dataf, Phylum == "Verrucomicrobia")

Actinobacteria<-filter(dataf, Phylum == "Actinobacteria")


```

Regress the relative abundance of the phyla against VTmax after the experiment. 
```{r}
#Firmicutes 
FirmLm<-lm(VTmax~Abundance, data = Firmicutes)
p1<-signif(summary(FirmLm)$coef[2,4], 5)

ProtLm<-lm(VTmax~Abundance, data = Proteobacteria)
p2<-signif(summary(ProtLm)$coef[2,4], 5)

BacLm<-lm(VTmax~Abundance, data = Bacteroidetes)
p3<-signif(summary(BacLm)$coef[2,4], 5)

ActinoLm<-lm(VTmax~Abundance, data = Actinobacteria)
p4<-signif(summary(ActinoLm)$coef[2,4], 5)

VerrucoLm<-lm(VTmax~Abundance, data = Verrucomicrobia)
p5<-signif(summary(VerrucoLm)$coef[2,4], 5)

```

Regress the relative abundance of the phyla against CTmin after the experiment. 

```{r}
FirmLm<-lm(CTmin~Abundance, data = Firmicutes)
p6<-signif(summary(FirmLm)$coef[2,4], 5)

ProtLm<-lm(CTmin~Abundance, data = Proteobacteria)
p7<-signif(summary(ProtLm)$coef[2,4], 5)

BacLm<-lm(CTmin~Abundance, data = Bacteroidetes)
p8<-signif(summary(BacLm)$coef[2,4], 5)

ActinoLm<-lm(CTmin~Abundance, data = Actinobacteria)
p9<-signif(summary(ActinoLm)$coef[2,4], 5)

VerrucoLm<-lm(CTmin~Abundance, data = Verrucomicrobia)
p10<-signif(summary(VerrucoLm)$coef[2,4], 5)
```

Regress the relative abundance of the phyla against Tpref mean, after 

```{r}
FirmLm<-lm(Tpref_mean~Abundance, data = Firmicutes)
p11<-signif(summary(FirmLm)$coef[2,4], 5)

ProtLm<-lm(Tpref_mean~Abundance, data = Proteobacteria)
p12<-signif(summary(ProtLm)$coef[2,4], 5)

BacLm<-lm(Tpref_mean~Abundance, data = Bacteroidetes)
p13<-signif(summary(BacLm)$coef[2,4], 5)

ActinoLm<-lm(Tpref_mean~Abundance, data = Actinobacteria)
p14<-signif(summary(ActinoLm)$coef[2,4], 5)

VerrucoLm<-lm(Tpref_mean~Abundance, data = Verrucomicrobia)
p15<-signif(summary(VerrucoLm)$coef[2,4], 5)
```

Extract the PCs of the unweighted unifrac results in order to regress against it. 
```{r}
#read in the pcoa results
unweighted_unifrac<-read_qza("./Climate-chamber/core-metrics-results/unweighted_unifrac_pcoa_results.qza")

#read in the metadata
metadata<-read_tsv("physiologyafterexperiment.txt")

#remove the second column of the metadata
metadata<-metadata[-1,]

#extract the data and the Vectors from the pcoa qza file. 
data<-unweighted_unifrac$data
data.vectors.t<-data$Vectors

# rename the first column to match the pcoa names 
rownames(metadata) <- NULL
colnames(metadata) <- c(names(metadata)) #to not write all the column names
colnames(metadata)[1] <- "SampleID"

#join the metadata and the pcoa
joined<-left_join(data.vectors.t, metadata)

#filter out the two samples that have NA values. 
joined.filtered<-filter(joined, joined$SampleID !='3V')
joined.filtered.1<-filter(joined.filtered, joined.filtered$SampleID !='3W')

#rename to final df 
finaldf<-joined.filtered.1

```

Regress VTmax after, CTmin after, and TPref mean against the first PC of the unweighted unifrac file 
```{r}
UU_VTM_PC1Lm<-lm(VTmax~PC1, data = finaldf)
p16<-signif(summary(UU_VTM_PC1Lm)$coef[2,4], 5)

UU_CTM_PC1Lm<-lm(CTmin~PC1, data = finaldf)
p17<-signif(summary(UU_CTM_PC1Lm)$coef[2,4], 5)

UU_TPM_PC1Lm<-lm(Tpref_mean~PC1, data = finaldf)
p18<-signif(summary(UU_TPM_PC1Lm)$coef[2,4], 5)

```

Regress VTmax after, CTmin after, and TPref mean against the second PC of the unweighted unifrac file 

```{r}
UU_VTM_PC2Lm<-lm(VTmax~PC2, data = finaldf)
p19<-signif(summary(UU_VTM_PC2Lm)$coef[2,4], 5)

UU_CTM_PC2Lm<-lm(CTmin~PC2, data = finaldf)
p20<-signif(summary(UU_CTM_PC2Lm)$coef[2,4], 5)

UU_TPM_PC2Lm<-lm(Tpref_mean~PC2, data = finaldf)
p21<-signif(summary(UU_TPM_PC2Lm)$coef[2,4], 5)

```


Extract the  of the weighted unifrac results in order to regress against it. 
```{r}
#read in the pcoa results
unweighted_unifrac<-read_qza("./Climate-chamber/core-metrics-results/weighted_unifrac_pcoa_results.qza")

#read in the metadata
metadata<-read_tsv("physiologyafterexperiment.txt")

#remove the second column of the metadata
metadata<-metadata[-1,]

#extract the data and the Vectors from the pcoa qza file. 
data<-unweighted_unifrac$data
data.vectors.t<-data$Vectors

# rename the first column to match the pcoa names 
rownames(metadata) <- NULL
colnames(metadata) <- c(names(metadata)) #to not write all the column names
colnames(metadata)[1] <- "SampleID"

#join the metadata and the pcoa
joined<-left_join(data.vectors.t, metadata)

#filter out the two samples that have NA values. 
joined.filtered<-filter(joined, joined$SampleID !='3V')
joined.filtered.1<-filter(joined.filtered, joined.filtered$SampleID !='3W')

#rename to final df 
finaldf<-joined.filtered.1

```
Regress VTmax after, CTmin after, and TPref mean against the first PC of the weighted unifrac file 
```{r}
WU_VTM_PC1Lm<-lm(VTmax~PC1, data = finaldf)
p22<-signif(summary(WU_VTM_PC1Lm)$coef[2,4], 5)

WU_CTM_PC1Lm<-lm(CTmin~PC1, data = finaldf)
p23<-signif(summary(WU_CTM_PC1Lm)$coef[2,4], 5)

WU_TPM_PC1Lm<-lm(Tpref_mean~PC1, data = finaldf)
p24<-signif(summary(WU_TPM_PC1Lm)$coef[2,4], 5)

```

Regress VTmax after, CTmin after, and TPref mean against the second PC of the weighted unifrac file 

```{r}
WU_VTM_PC2Lm<-lm(VTmax~PC2, data = finaldf)
p25<-signif(summary(WU_VTM_PC2Lm)$coef[2,4], 5)

WU_CTM_PC2Lm<-lm(CTmin~PC2, data = finaldf)
p26<-signif(summary(WU_CTM_PC2Lm)$coef[2,4], 5)

WU_TPM_PC2Lm<-lm(Tpref_mean~PC2, data = finaldf)
p27<-signif(summary(WU_TPM_PC2Lm)$coef[2,4], 5)

```


Extract the PCs of the Jaccard results in order to regress against it. 
```{r}
#read in the pcoa results
axes<-read_qza("./Climate-chamber/core-metrics-results/jaccard_pcoa_results.qza")

#read in the metadata
metadata<-read_tsv("physiologyafterexperiment.txt")

#remove the second column of the metadata
metadata<-metadata[-1,]

#extract the data and the Vectors from the pcoa qza file. 
data<-axes$data
data.vectors.t<-data$Vectors

# rename the first column to match the pcoa names 
rownames(metadata) <- NULL
colnames(metadata) <- c(names(metadata)) #to not write all the column names
colnames(metadata)[1] <- "SampleID"

#join the metadata and the pcoa
joined<-left_join(data.vectors.t, metadata)

#filter out the two samples that have NA values. 
joined.filtered<-filter(joined, joined$SampleID !='3V')
joined.filtered.1<-filter(joined.filtered, joined.filtered$SampleID !='3W')

#rename to final df 
finaldf<-joined.filtered.1

```
Regress VTmax after, CTmin after, and TPref mean against the first PC of the Jaccard file 
```{r}
J_VTM_PC1Lm<-lm(VTmax~PC1, data = finaldf)
p28<-signif(summary(J_VTM_PC1Lm)$coef[2,4], 5)

J_CTM_PC1Lm<-lm(CTmin~PC1, data = finaldf)
p29<-signif(summary(J_CTM_PC1Lm)$coef[2,4], 5)

J_TPM_PC1Lm<-lm(Tpref_mean~PC1, data = finaldf)
p30<-signif(summary(J_TPM_PC1Lm)$coef[2,4], 5)

```

Regress VTmax after, CTmin after, and TPref mean against the second PC of the Jaccard file 

```{r}
J_VTM_PC2Lm<-lm(VTmax~PC2, data = finaldf)
p31<-signif(summary(J_VTM_PC2Lm)$coef[2,4], 5)

J_CTM_PC2Lm<-lm(CTmin~PC2, data = finaldf)
p32<-signif(summary(J_CTM_PC2Lm)$coef[2,4], 5)

J_TPM_PC2Lm<-lm(Tpref_mean~PC2, data = finaldf)
p33<-signif(summary(J_TPM_PC2Lm)$coef[2,4], 5)

```


Extract the PCs of the bray curtis results in order to regress against them 
```{r}
#read in the pcoa results
axes<-read_qza("./Climate-chamber/core-metrics-results/bray_curtis_pcoa_results.qza")

#read in the metadata
metadata<-read_tsv("physiologyafterexperiment.txt")

#remove the second column of the metadata
metadata<-metadata[-1,]

#extract the data and the Vectors from the pcoa qza file. 
data<-axes$data
data.vectors.t<-data$Vectors

# rename the first column to match the pcoa names 
rownames(metadata) <- NULL
colnames(metadata) <- c(names(metadata)) #to not write all the column names
colnames(metadata)[1] <- "SampleID"

#join the metadata and the pcoa
joined<-left_join(data.vectors.t, metadata)

#filter out the two samples that have NA values. 
joined.filtered<-filter(joined, joined$SampleID !='3V')
joined.filtered.1<-filter(joined.filtered, joined.filtered$SampleID !='3W')

#rename to final df 
finaldf<-joined.filtered.1

```
Regress VTmax after, CTmin after, and TPref mean against the first PC of the Jaccard file 
```{r}
BC_VTM_PC1Lm<-lm(VTmax~PC1, data = finaldf)
p40<-signif(summary(BC_VTM_PC1Lm)$coef[2,4], 5)

BC_CTM_PC1Lm<-lm(CTmin~PC1, data = finaldf)
p41<-signif(summary(BC_CTM_PC1Lm)$coef[2,4], 5)

BC_TPM_PC1Lm<-lm(Tpref_mean~PC1, data = finaldf)
p42<-signif(summary(BC_TPM_PC1Lm)$coef[2,4], 5)

```

Regress VTmax after, CTmin after, and TPref mean against the second PC of the Jaccard file 

```{r}
BC_VTM_PC2Lm<-lm(VTmax~PC2, data = finaldf)
p43<-signif(summary(BC_VTM_PC2Lm)$coef[2,4], 5)

BC_CTM_PC2Lm<-lm(CTmin~PC2, data = finaldf)
p44<-signif(summary(BC_CTM_PC2Lm)$coef[2,4], 5)

BC_TPM_PC2Lm<-lm(Tpref_mean~PC2, data = finaldf)
p45<-signif(summary(BC_TPM_PC2Lm)$coef[2,4], 5)

```

Correct all P-values based on multiple comparisons. 
```{r}
list<-c(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,p13,p14,p15,p16,p17,p18,p19,p20,
     p21,p22,p23,p24,p25,p26,p27,p28,p29,p30,p31,p32,p33,p34,p35,p36,p37,p38,
     p39,p40,p41,p42,p43,p44,p45)

adjusted<-p.adjust(list, method = "bonferroni", n = length(list))

p.values<-cbind(list, adjusted) %>%
  as.data.frame()

#write_csv(p.values, "physiology-results.csv")

```





